\section{Appendix}

\textbf{The questionnaire}

\begin{enumerate}[label*=\arabic*.]
  \item \textbf{General Part}
% \begin{itemize}
 \begin{enumerate}[label*=\arabic*.]
   \item Why/How have you engaged in the project for the first time?
   \item How long have you been participating?
   \item What exactly are you doing/have you done within Freifunk?
   \item Are you actively participating in organising the community work: meetings, etc.?
   \item Which aspects of the project are cool? Which are not so cool?
   \item What is your background?
% \end{itemize}
 \end{enumerate}

\item \textbf{More specific:} Are following factors of significance for your motivation?
%\begin{itemize}
\begin{enumerate}[label*=\arabic*.]
  \item Freifunk is technically challenging and it's fun to work on the project.
  \item I want to learn the technical aspects of how wifi/mesh networks/etc. work.
  \item I wanted to learn more about the legal framework within which the project operates.
  \item I want to have a free communication infrastructure in the hands of civil society, because I think it's not a good idea that this is controlled only by the state and a couple of business players.
  \item I started to participate because commercial providers didn't have any interest of providing me with an efficient Internet connection.
  \item I care for net neutrality.
  \item I'm providing professional services related to Freifunk: I receive money for hardware installations, maintenance, etc.
  \item I want to work in this field later, so my participation in Freifunk is an excellent opportunity to polish my technical skills, make new contacts and become visible for potential employers.
%\end{itemize}
\end{enumerate}

\end{enumerate}


 \begin{comment}
 * Warum machst du Freifunk? --> übergeordnete Frage
1.1 Warum/Wie bist du ins Projekt eingestiegen?
1.2 Wie lange machst du das schon?
1.3 Was genau machst du da/hast du bis jetzt gemacht?
1.4 Wie aktiv nimmst du an Orga teil?
1.5 Was findest du an dem Projekt cool? Was findest du weniger cool?
1.6 hintergrund?
\end{comment}

\begin{comment}
Spielen folgende Punkte ne Rolle für deine Motivation:
2.1 Freifunk ist technisch spannend/anspruchsvoll und es macht Spaß, sich damit auseinander zu setzen
2.2 Ich will mir die technischen Fähigkeiten aneingnen, wie WLAN/Mesh-Netzwerke/... funktionieren
2.3 Ich will mich mit den gesetzlichen Rahmenbedinungen auseinandersetzen
2.4 Ich will eine freie, community-betriebene Kommunikationsinfrastruktur (mit)schaffen, weil ich denke, es ist Scheiße, dass alles in den Händen von paar Konzernen ist
2.5 Ich hab damit angefangen, weil die kommerziellen Providern keine Lust hatten, mich anzuschließen, weil es sich für sie nicht gelohnt hat
2.6 Mir ist Netzneutralität wichtig
2.7 Ich mach das mit Freifunk beruflich: krieg Geld für Knoten einrichten, Wartung, ...
2.8 Ich mach/will machen was auf dem Gebiet beruflich und das ist die Gelegenheit, meine Fähigkeiten aufzupollieren/Kontakte knüpfen/für Arbeitgeber_innen sichtbar werden
\end{comment}

